Meta RL analyses

### 2019-02-08 writing/embedded_meta_learning.tex
Some more math

\section{Introduction}
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep{Lake2017}. One fundamental reason for this is the separation of knowledge within deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between these. That is, neural networks knowledge about \textbf{what} is being computed is \textbf{inaccessible} by the computations. By contrast, humans are frequently consciously aware of both our representations of input, the things we can do with them, and the relationships between data and tasks. This is true for domains from game playing to natural language understanding. \par
One of the abilities that this grants humans is rapid adaptation of our behavior. The problem of rapid learning has been partially addressed by meta-learning systems \citep{}. However, humans can take rapidly acquired knowledge and flexibly alter their behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance \textbf{without any retraining} (i.e. zero-shot). Deep learning systems at present lack this representational flexiblity. \par
In this paper, we propose a new class of architectures which essentially take a \textbf{functional perspective} on meta-learning. By treating tasks as functions, we can conceptually think of both data \textbf{and} learned functions as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing task demands. We suggest that this ability will be key to building more intelligent and flexible deep learning systems. \par
In particular, we propose a meta-learning architecture which embeds data (including natural language inputs) and tasks in a shared space, and meta-learns to execute mappings over this space. We demonstrate the greater flexibility of this approach.
\section{Method}
\subsection{Architecture}
The essential idea of the architecture is that we separate the system into two parts: 
\begin{enumerate}
\item Domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$
\item A meta-learning system which:
    \begin{enumerate}
    \item Embeds tasks into the shared embedding space $Z$.
    \item Learns to perform tasks in accordance with their embeddings.
    \end{enumerate}
\end{enumerate}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, ideally if it is trained on a set of such mappings, it will be able to generalize these to new instances drawn from the same data/task distribution. For instances that fall outside its data distribution or for optimal performance it may require some retraining. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relateively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow. \par

@@

\end{tikzpicture}
\caption{Schematic of our general architecure. Blocks with solid edges denote deep networks with learnable parameters, dashed edges represent inputs, outputs, embeddings, etc., and $f$ is a deep network with parameters specified by $\mathcal{H}$.} \label{architecture_fig}
\end{figure}
We take a \textbf{functional} perspective on learning. A datum can be represented by a constant function which outputs it. (For example, each point in the latent space of an autoencoder can be thought of this way.) This allows us to interpret model inputs or outputs as functions. \par
We can then interpret most machine learning tasks as a mapping of functions to functions. These functions could represent data\footnote{Where ``data'' is a quite flexible term. The approach is relatively agnostic to whether the learning is supervised or reinforcement learning, e.g.}, or they could be functions that operate on functions themselves. Under this perspective, learning tasks and learning to flexibly map between tasks are the same type of problem. \par
More specifically, we embed inputs, targets, and mappings into a shared representational space $Z$. Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Outputs are decoded from the representational space by a deep network $\mathcal{O}: Z \rightarrow \text{output}$. Targets are encoded by a deep network $\mathcal{T}: \text{targets} \rightarrow Z$. (Targets do not necessarily need to be output-like, e.g. in our RL tasks, we use (action, outcome) tuples as ``targets.'') \par
Given this, the task of mapping inputs to outputs can be framed as trying to find a transformation of the representational space that takes the (embedded) inputs from the training set to the (embedded) targets. These transformations are performed by a system with the following components:  
\begin{itemize}
\item $\mathcal{M}: \{(Z, Z), ...\} \rightarrow Z $ -- the meta network, which takes a set of (input embedding, target embedding) pairs and produces a function embedding. 
\item $\mathcal{H}: Z \rightarrow \text{parameters}$ -- the hyper network, which takes a function embedding and produces a set of parameters. 
\item $f: Z \rightarrow Z$ -- the transformation, implemented by a deep network with parameters specified by $\mathcal{H}$.
\end{itemize}
See Fig. \ref{architecture_fig} for a schematic of the architecture. \par

\subsection{Operation}
A basic forward pass through the system might look as follows.
\begin{enumerate}
\item A training dataset of (input, target) pairs is embedded by $\mathcal{I}$ and $\mathcal{T}$ to produce a set of paired embeddings. Another set of (possibly unlabeled) inputs is provided and embedded.
\item The meta network $\mathcal{M}$ maps the set of embedded (input, target) pairs to a function embedding.
\item The hyper network $\mathcal{H}$ maps the function embedding to parameters for $f$, which is used to transform the second set of inputs to a set of output embeddings.
\item The output embeddings are decoded by $\mathcal{O}$ to produce a set of outputs.
\end{enumerate}
This system can be trained end-to-end if labels are provided for the second set of inputs. In particular, suppose we have some dataset of input, target pairs ($D_1 = \{(x_0, y_0), ...\}$), and some loss function $\mathcal{L}(y, \hat{y})$ defined on a single target output $y$ and actual model output $\hat{y}$, for some input $x$. We define our total loss computed on some dataset $D_2$ as:
$$\mathbb{E}_{(x, y)\in {D}_2} \left[ \mathcal{L}\left(y, \mathcal{O}\left(f_{D_1}\left(\mathcal{I} \left(x\right)\right) \right)\right)\right]$$
where $f_{D_1}$ is the transformation the meta-learner guesses for the training dataset $D_1$:
$$f_{D_1} \text{ is parameterized by } \mathcal{H}\left(\mathcal{M}\left( \left\{\left(\mathcal{I}\left(x_0\right), \mathcal{T}\left(y_0\right) \right), \left(\mathcal{I}\left(x_1\right), \mathcal{T}\left(y_1\right) \right), ... \right\}\right)\right)$$

@@

\subsection{A reinforcment learning setting: simple card games}
Our next case study was in the setting of simple card games, played with two suits, and 4 values per suit. In our setup, each hand in a game has a win probability (proportional to how it ranks against all other possible hands). The agent is dealt a hand, and then has to choose to bet 0, 1, or 2. If it wins, it is rewarded the value of the bet, if it loses, it receives the negative of the value. This doesn't require long term planning, but is nevertheless an RL rather than a supervised problem, since the feedback is stochastic. We considered a variety of games:
\begin{itemize}
\item \textbf{High card:} Highest card wins.
\item \textbf{Pairs and high:} Same as high card, except pairs are more valuable, and same suit pairs are even more valuable.
\item \textbf{Straight flush:} Most valuable is adjacent numbers in same suit, i.e. 4 and 3 in most valuable suit wins every time (royal flush).
\item \textbf{Match:} the hand with cards that differ least in value (suit counts as 0.5 pt difference) wins.
\item \textbf{Sum under:} The hand's value increases with the sum of the cards until it crosses 5, at which point the player ``goes bust,'' and the value becomes negative. 
\end{itemize}
These games are simplified analogs of various card games, including poker and blackjack. We also considered three binary options that could be altered to produce variants of these games:
\begin{itemize}
\item \textbf{Losers:} Try to lose instead of winning! Reverses the ranking of hands.
\item \textbf{Suits rule:} Instead of suits being less important than values, they are more important (essentially flipping the role of suit and value in most games).
\item \textbf{Switch suit:} Switches which of the suits is more valuable.
\end{itemize}
Any combination of these options can be applied to any of the 5 games, yielding 40 possible games. \par
In order to switch the system to an RL meta learning system, we made a very simple change. Instead of providing the system with (input, target) tuples to embed, we provided it with (state, action, reward) tuples, and trained it via a DQN-like approach \citep{} to predict rewards for each bet in each state. The hand is explicitly provided to the network for each example, but which game is being played is implicitly captured in the training examples, without any explicit cues. That is, the system must learn to play directly from seeing a set of (state, action, reward) tuples. \par   
We can train the system to identify the various types of games and options, and to make meta-mappings, for example switching from trying to win a game to trying to lose. \par

>>>>

\section{Introduction}
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep{Lake2017}. One fundamental reason for this is the separation of knowledge within deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between these. That is, neural networks knowledge about \textbf{what} is being computed is \textbf{inaccessible} by the computations. By contrast, humans are frequently consciously aware of both our representations of input, the things we can do with them, and the relationships between data and tasks. This is true for domains from game playing to natural language understanding. \par
One of the abilities that this grants humans is rapid adaptation of our behavior. The problem of rapid learning has been partially addressed by meta-learning systems \citep{}. However, humans can take rapidly acquired knowledge and flexibly alter their behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance \textbf{without any retraining} (i.e. zero-shot). Deep learning systems at present lack this representational flexiblity. \par
In this paper, we propose a new class of architectures which essentially take a \textbf{functional perspective} on meta-learning. By treating tasks as functions, we can conceptually think of both data \textbf{and} learned functions as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing task demands. We suggest that this ability will be key to building more intelligent and flexible deep learning systems. \par
In particular, we propose a meta-learning architecture which embeds data (including natural language inputs) and tasks in a shared space, and meta-learns to execute mappings over this space. We demonstrate the greater flexibility of this approach.
\section{Method}
\subsection{Architecture}
The essential idea of the architecture is that we separate the system into two parts: 
\begin{enumerate}
\item Domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$
\item A meta-learning system which:
    \begin{enumerate}
    \item Embeds tasks into the shared embedding space $Z$.
    \item Learns to perform tasks in accordance with their embeddings.
    \end{enumerate}
\end{enumerate}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, ideally if it is trained on a braod enough set of such mappings, it will be able to generalize these to new instances drawn from the same data/task distribution. For instances that fall outside its data distribution or for optimal performance it may require some retraining. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relateively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow. \par

@@

\end{tikzpicture}
\caption{Schematic of our general architecure. Blocks with solid edges denote deep networks with learnable parameters, dashed edges represent inputs, outputs, embeddings, etc., and $f$ is a deep network with parameters specified by $\mathcal{H}$.} \label{architecture_fig}
\end{figure}
More formally, we take a \textbf{functional} perspective on learning. A datum can be represented by a constant function which outputs it. (For example, each point in the latent space of an autoencoder can be thought of this way.) This allows us to interpret model inputs or outputs as functions. \par
We can then interpret most machine learning tasks as a mapping of functions to functions. These functions could represent data\footnote{Where ``data'' is a quite flexible term. The approach is relatively agnostic to whether the learning is supervised or reinforcement learning, whether inputs are images or natural language, etc.}, or they could be functions that operate on functions themselves. Under this perspective, learning tasks and learning to flexibly map between tasks, or learning to map from language to tasks, are all the same type of problem. \par
Specifically, we embed inputs, targets, and mappings into a shared representational space $Z$. Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Outputs are decoded from the representational space by a deep network $\mathcal{O}: Z \rightarrow \text{output}$. Targets are encoded by a deep network $\mathcal{T}: \text{targets} \rightarrow Z$. (Targets do not necessarily need to be output-like, e.g. in our RL tasks, we use (action, outcome) tuples as ``targets.'') \par
Given this, the task of mapping inputs to outputs can be framed as trying to find a transformation of the representational space that takes the (embedded) inputs from the training set to the (embedded) targets. These transformations are performed by a system with the following components:  
\begin{itemize}
\item $\mathcal{M}: \{(Z, Z), ...\} \rightarrow Z $ -- the meta network, which takes a set of (input embedding, target embedding) pairs and produces a function embedding. 
\item $\mathcal{H}: Z \rightarrow \text{parameters}$ -- the hyper network, which takes a function embedding and produces a set of parameters. 
\item $f: Z \rightarrow Z$ -- the transformation, implemented by a deep network with parameters specified by $\mathcal{H}$.
\end{itemize}
See Fig. \ref{architecture_fig} for a schematic of the architecture. \par

\subsection{Operation}
A basic forward pass through the system might look as follows.
\begin{enumerate}
\item A training dataset of (input, target) pairs is embedded by $\mathcal{I}$ and $\mathcal{T}$ to produce a set of paired embeddings. Another set of (possibly unlabeled) inputs is provided and embedded.
\item The meta network $\mathcal{M}$ maps the set of embedded (input, target) pairs to a function embedding.
\item The hyper network $\mathcal{H}$ maps the function embedding to parameters for $f$, which is used to transform the second set of inputs to a set of output embeddings.
\item The output embeddings are decoded by $\mathcal{O}$ to produce a set of outputs.
\end{enumerate}
To write this explicitly, suppose we have some dataset of input, target pairs ($D_1 = \{(x_0, y_0), ...\}$), and some input $x$ for which we wish to generate a predicted output $\hat{y}$. This output would be generated as follows: 
$$\hat{y} = \mathcal{O}\left(f_{D_1}\left(\mathcal{I} \left(x\right)\right) \right)$$
where $f_{D_1}$ is the transformation the meta-learner guesses for the training dataset $D_1$:
$$f_{D_1} \text{ is parameterized by } \mathcal{H}\left(\mathcal{M}\left( \left\{\left(\mathcal{I}\left(x_0\right), \mathcal{T}\left(y_0\right) \right), \left(\mathcal{I}\left(x_1\right), \mathcal{T}\left(y_1\right) \right), ... \right\}\right)\right)$$
This system can be trained end-to-end if labels are provided for a second set of inputs. In particular, suppose we have some loss function $\mathcal{L}(y, \hat{y})$ defined on a single target output $y$ and actual model output $\hat{y}$, for some input $x$. We define our total loss computed on some dataset $D_2$ as:
$$\mathbb{E}_{(x, y)\in {D}_2} \left[ \mathcal{L}\left(y, \mathcal{O}\left(f_{D_1}\left(\mathcal{I} \left(x\right)\right) \right)\right)\right]$$
More generally, suppose we have some input which is already embedded in the representation space $z_{in} \in Z$, and an embedded dataset $D_Z$ of (embedding, target embedding) pairs $\{(z_{in,0}, z_{out,0})\}$. Then we can generate an output $\hat{z}_{out} \in Z$ as:  
$$\hat{z}_{out} = f_{D_Z}(z_{in}) \qquad \text{where } f_{D_Z} \text{ is parameterized by } \mathcal{H}\left(\mathcal{M}\left(D_Z\right)\right)$$
This output can then be appropriately dispatched depedning on the task at hand. For example, if the $z_{in,i}$ are the system's embeddings for trying to win various games, and the $z_{out,i}$ are the corresponding embeddings for trying to lose those games, then $z_{out}$ could be interpreted as the system's guess at a losing strategy for the game embedded as $z_{in}$, and then could be used to play that game. We could then evaluate performance by how well the system did at losing.

@@

\subsection{A stochastic learning setting: simple card games}
Our next case study was in the setting of simple card games, played with two suits, and 4 values per suit. In our setup, each hand in a game has a win probability (proportional to how it ranks against all other possible hands). The agent is dealt a hand, and then has to choose to bet 0, 1, or 2. If it wins, it is rewarded the value of the bet, if it loses, it receives the negative of the value. This doesn't require long term planning, but is nevertheless not precisely a supervised problem, since the feedback is stochastic. We considered a variety of games:
\begin{itemize}
\item \textbf{High card:} Highest card wins.
\item \textbf{Pairs and high:} Same as high card, except pairs are more valuable, and same suit pairs are even more valuable.
\item \textbf{Straight flush:} Most valuable is adjacent numbers in same suit, i.e. 4 and 3 in most valuable suit wins every time (royal flush).
\item \textbf{Match:} the hand with cards that differ least in value (suit counts as 0.5 pt difference) wins.
\item \textbf{Sum under:} The hand's value increases with the sum of the cards until it crosses 5, at which point the player ``goes bust,'' and the value becomes negative. 
\end{itemize}
These games are simplified analogs of various card games, including poker and blackjack. We also considered three binary options that could be altered to produce variants of these games:
\begin{itemize}
\item \textbf{Losers:} Try to lose instead of winning! Reverses the ranking of hands.
\item \textbf{Suits rule:} Instead of suits being less important than values, they are more important (essentially flipping the role of suit and value in most games).
\item \textbf{Switch suit:} Switches which of the suits is more valuable.
\end{itemize}
Any combination of these options can be applied to any of the 5 games, yielding 40 possible games. \par
In order to switch the system to an RL meta learning system, we made a very simple change. Instead of providing the system with (input, target) tuples to embed, we provided it with (state, action, reward) tuples, and trained it via a DQN-like approach \citep{} to predict rewards for each bet in each state. The hand is explicitly provided to the network for each example, but which game is being played is implicitly captured in the training examples, without any explicit cues. That is, the system must learn to play directly from seeing a set of (state, action, reward) tuples. \par   
We can train the system to identify the various types of games and options, and to make meta-mappings, for example switching from trying to win a game to trying to lose. \par
